---
title: "Analyzing trends in Canadian glacier mass"
author: "Anton Antonov"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The data used in this study is taken from <http://open.canada.ca/en/open-data>[^1].

[^1]: Here is the [direct link](http://open.canada.ca/data/en/dataset/b52664cf-bfd9-49ad-849a-cb88c92553b9) to data download.

```{r, echo=FALSE}
glacier <- read.csv("glacier.csv", na.strings = "..", comment.char = "#")
```

The data set contains `r nrow(glacier)` measurements of `r nlevels(glacier$GEO)` Canadian glacier mass balance, collected from `r min(glacier$Ref_Date)` till `r max(glacier$Ref_Date)`. Namely, the file includes these glaciers:

```{r, echo=FALSE}
levels(glacier$GEO)
```

## Hypothesis

We are interested in finding out whether there is a statistically significant change in mass balance over the observed time period. For these purposes we use **R** (version `r getRversion()`) and an appropriate statistical test called _t-test_: $$t = \frac{\overline{x} - \mu_0}{s/\sqrt{n}}.$$

The workflow is as follows:

1. Read the file and run data validation;
2. Run t-test for each glacier and collect p-values;
3. Support the evidence with
    + a table of results;
    + a plot that could help compare different glaciers.
