---
title: "Analyzing trends in Canadian glacier mass"
author: "Anton Antonov"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(xtable)
library(ggplot2)
options(scipen = 0, digits = 3)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Introduction

The data used in this study is taken from <http://open.canada.ca/en/open-data>[^1].

[^1]: Here is the [direct link](http://open.canada.ca/data/en/dataset/b52664cf-bfd9-49ad-849a-cb88c92553b9) to data download.

```{r}
glacier <- read.csv("glacier.csv", na.strings = "..", comment.char = "#")
```

The data set contains `r nrow(glacier)` measurements of `r nlevels(glacier$GEO)` Canadian glacier mass balance, collected from `r min(glacier$Ref_Date)` till `r max(glacier$Ref_Date)`. Namely, the file includes these glaciers:

```{r, comment=""}
levels(glacier$GEO)
```

## Hypothesis

We are interested in finding out whether there is a statistically significant change in mass balance over the observed time period. For these purposes we use **R** (version `r getRversion()`) and an appropriate statistical test called _t-test_: $$t = \frac{\overline{x} - \mu_0}{s/\sqrt{n}}.$$

The workflow is as follows:

1. Read the file and run data validation;
2. Run t-test for each glacier and collect p-values;
3. Support the evidence with
    + a table of results;
    + a plot that could help compare different glaciers.
    
## Analysis

```{r analysis, cache=TRUE}
glacier <- read.csv("glacier.csv", na.strings = "..", comment.char = "#")
glacier <- glacier %>%
  select(Ref_Date, GEO, MEASURE, Value) %>% 
  filter(MEASURE == "Annual mass balance") %>% 
  separate(GEO, c("Name", "Location"), sep = " - ")

g1 <- glacier %>%
  group_by(Name) %>% 
  summarise(YearsObserved = n(), 
            MeanChange = mean(Value, na.rm = T), 
            WorstChange = min(Value, na.rm = T),
            WorstYear = Ref_Date[which.min(Value)])

g2 <- glacier %>% 
  group_by(Name) %>% 
  do({
    tt <- t.test(.$Value, alternative = "less", mu = 0, conf.level = 0.99)
    data.frame(PValue = tt$p.value, 
               ConfidenceLimit = tt$conf.int[2])
    })

left_join(g1, g2, by = "Name") %>% 
  knitr::kable(caption = "Descriptive statistics and confidence intervals", 
               digits = c(0, 0, 2, 0, 0, 10, 2))
```

```{r ggplot, fig.width=8, fig.height=6}
ggplot(glacier, aes(Ref_Date, Value)) + 
  geom_line() + 
  geom_hline(data = g1, aes(yintercept = MeanChange), 
             color = "red", linetype = "dashed", alpha = 0.8) + 
  facet_wrap(~Name, nrow = 2)
```


